---
title: "Your A4 Report Title (TODO: Revise title)"
author: "Jenna Moon-Earle"
date: '2022-11-01'
output: html_document
---

```{r eval=TRUE, echo=FALSE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
The justice system of America is said to be fair, but there are still manifestations of racism in these systems. Observing incarceration trends can answer the questions of how prevalent racism is in the prison systems, if and how it has progressed over the years, and other forms of inequalities there are in the systems.  

## Introduction 
[TODO]

## Data Summary 
Question 1: What is the percentage of black individuals in jail compared to total jail population in the year 2018? 

The percentage of black individuals in jail is **`r black_jail_prop`**%. 

Question 2: What year has the highest black population in jail? 

**`r highest_black_pop_year`** had the highest black population in jail. 

Question 3: What county has the highest population of black and latinx individuals in jail in 2018?

**`r county_highest_black_pop`** has the highest population of black individuals in jail.

**`r county_highest_latinx_pop` has the highest population of latinx individuals in jail.

## U.S. Jail Population Growth
```{r fig.height = 8, fig.width = 8, fig.align = "center", echo = FALSE}
plot_jail_pop_for_us()
```

## State Jail Population Growth 
```{r fig.height = 8, fig.width= 8, fig.align = "center", echo = FALSE}
states_pop_plot
```

## Section 5
```{r fig.height = 8, fig.width= 8, fig.align = "center", echo = FALSE}
section5plot()
```

## Section 6
[TODO]
